Arc Blanc: a real time ocean simulation framework

### 2024-11-12 paper.tex
Fixed author name

While there are many articles describing different parts of ocean simulation, the literature lacks papers that bring all the pieces together.
This article proposes a fully described real-time framework designed to simulate the free surface of the ocean and the interactions between solid objects and the surrounding fluid in real time (\cf Figure \ref{fig:overviewFramework}).
This new framework incorporates three different methods from the literature:
\begin{itemize}
	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
	\item The simulation of the alteration of a free surface by a solid, inspired by Coords and Staadt works (2009) \cite{cordsRealTimeOpenWater2009}.
\end{itemize}
To bring together these previous works, this new framework proposes two new contributions:
\begin{itemize}
	\item An algorithm to compute the ocean velocity at any depth from the simulation of the free surface. This velocity is applied to the fluid to solid forces.
	\item An algorithm to compute the input of the Coords and Staadt method more realistically from the free surface.
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Solid to fluid coupling}
\label{subsubsec:solidToFluidCoupling}
The simulation of the surface of water in contact with a solid body has been an active research topic for the last twenty years.
Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the FDM to simulate waves on the free surface.
This method was improved by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
Tessendorf (2004) \cite{tessendorfInteractiveWaterSurface2004} proposed a convolutional method.
He extended his method by using an exponential solver \cite{tessendorfEWaveUsingExponential2014}.
Yuksel (2010) \cite{yukselRealtimeWaterWaves2010} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
Chentanez and Müller (2010) \cite{chentanezRealTimeSimulationLarge2010} created a hybrid solver combining Eulerian and Lagrangian approaches.
Canabal  \etal (2016) \cite{canabalDispersionKernelsWater2016} improved the physical accuracy of the convolutional method by introducing a dispersion kernel.
Jeschke and Woktan (2017) \cite{jeschkeWaterWavePackets2017} took the work of C. Yuskel a bit further by introducing the notion of a wave packet, which carries information about an entire wave train.
One year later, Jeschke  \etal (2018) \cite{jeschkeWaterSurfaceWavelets2018} proposed to use a wavelet transformation that discretizes the wave amplitudes as a function of space, frequency and direction.
Finally, Schreck  \etal (2019) \cite{schreckFundamentalSolutionsWater2019} based their work on the method of fundamental solutions to generate ambient waves on a large scale domain interacting with complex boundaries.

@@

%-------------------------------------------------------------------------
\section{Solid-to-fluid action}
\label{sec:solidToFluidAction}
\Framework relies on the method developed by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to simulate the action of the solid on the fluid.
The simulation of this action is 3-dimensional by nature.
However, it has to be simplified to satisfy the third requirement of Section~\ref{sec:couplingSolidAndFluid}, concerning computation time performance.
As for fluid simulation in Section~\ref{sec:oceanWaves}, \framework assumes that the interaction from a solid to the fluid can be simulated as a 2D height field.
Furthermore, it considers a single frequency for the waves generated by a solid $M$.
These specific waves are called interactive waves.
These approximations may seem too strong, but as mentioned above, they are necessary to satisfy the requirements defined in Section~\ref{sec:couplingSolidAndFluid}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Grid translation}
\label{subsec:gridTranslation}
The waves generated by a given body $M$ are defined in its local coordinates system.
Hence, when $M$ moves, the waves have to move accordingly as done by Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
A first solution consists of applying the translation to the grid at times $t$ and $t-dt$ and then of using the integration scheme of Equation (\ref{eq:fdm_scheme}).
Nevertheless, these time-consuming shifts can be avoided by using the following domain translation:
\begin{equation}
	\begin{cases}
		k = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t)}{dx}\right\rfloor    \\
		l = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t)}{dx}\right\rfloor    \\
		o = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t-dt)}{dx}\right\rfloor \\
		p = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t-dt)}{dx}\right\rfloor
	\end{cases}
\end{equation}
where $\mathbf p(t)=\left[p_x(t)\ p_z(t)\right]^T$ is the position of the mesh $M$ at time $t$, $k$ and $l$ (respectively $o$ and $p$) are the shifted indices defined by the translation of $M$ at times $t+dt$ from times $t$ (respectively from times $t-dt$).
Then, the integration scheme can be rewritten as follows:
\begin{equation}
	\label{eq:fdm_scheme_corrected}
	h_{i,j}^{n+1} = d^n\left(a\left(h_{k+1,l}^n+h_{k-1,l}^n+h_{k,l+1}^n+h_{k,l-1}^n- 4 h_{k,l}^n\right) + 2 h_{k,l}^n - h_{o,p}^{n-1}\right)
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Evaluate the height of vertices in the mask}
\label{subsubsec:heightMask}
Knowing all the vertices in the mask, their height must be given as input of the FDM.
\Framework uses a simple model that tries to reproduce parts of the idea of Coords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
By construction of the grid $Z$, $M$ has is bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
Hence, $h$ is made such that from the vertex at position $\mathbf x$, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of $M$.
Moreover, the boat forms a $V$ shape, with its center a bit below the extremity of the side.

>>>>

While there are many articles describing different parts of ocean simulation, the literature lacks papers that bring all the pieces together.
This article proposes a fully described real-time framework designed to simulate the free surface of the ocean and the interactions between solid objects and the surrounding fluid in real time (\cf Figure \ref{fig:overviewFramework}).
This new framework incorporates three different methods from the literature:
\begin{itemize}
	\item The simulation of the free surface of the ocean, initially developed by Tessendorf (2004) \cite{tessendorfSimulatingOceanWater1999} and its improvements along the years \cite{horvathEmpiricalDirectionalWave2015,tessendorfGilliganPrototypeFramework2017}.
	\item The estimation of the forces applied by the fluid on a solid mesh, mixing together different techniques \cite{yukselRealtimeWaterWaves2010,kellomakiRigidBodyInteraction2014}.
	\item The simulation of the alteration of a free surface by a solid, inspired by Cords and Staadt works (2009) \cite{cordsRealTimeOpenWater2009}.
\end{itemize}
To bring together these previous works, this new framework proposes two new contributions:
\begin{itemize}
	\item An algorithm to compute the ocean velocity at any depth from the simulation of the free surface. This velocity is applied to the fluid to solid forces.
	\item An algorithm to compute the input of the Cords and Staadt method more realistically from the free surface.
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Solid to fluid coupling}
\label{subsubsec:solidToFluidCoupling}
The simulation of the surface of water in contact with a solid body has been an active research topic for the last twenty years.
Gomez (2000) \cite{gomezInteractiveSimulationWater2000} solved the wave equation using the FDM to simulate waves on the free surface.
This method was improved by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to work with moving objects.
Tessendorf (2004) \cite{tessendorfInteractiveWaterSurface2004} proposed a convolutional method.
He extended his method by using an exponential solver \cite{tessendorfEWaveUsingExponential2014}.
Yuksel (2010) \cite{yukselRealtimeWaterWaves2010} used a Lagrangian paradigm to model the waves as particles, which he called wave-particles.
Chentanez and Müller (2010) \cite{chentanezRealTimeSimulationLarge2010} created a hybrid solver combining Eulerian and Lagrangian approaches.
Canabal  \etal (2016) \cite{canabalDispersionKernelsWater2016} improved the physical accuracy of the convolutional method by introducing a dispersion kernel.
Jeschke and Woktan (2017) \cite{jeschkeWaterWavePackets2017} took the work of C. Yuskel a bit further by introducing the notion of a wave packet, which carries information about an entire wave train.
One year later, Jeschke  \etal (2018) \cite{jeschkeWaterSurfaceWavelets2018} proposed to use a wavelet transformation that discretizes the wave amplitudes as a function of space, frequency and direction.
Finally, Schreck  \etal (2019) \cite{schreckFundamentalSolutionsWater2019} based their work on the method of fundamental solutions to generate ambient waves on a large scale domain interacting with complex boundaries.

@@

%-------------------------------------------------------------------------
\section{Solid-to-fluid action}
\label{sec:solidToFluidAction}
\Framework relies on the method developed by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009} to simulate the action of the solid on the fluid.
The simulation of this action is 3-dimensional by nature.
However, it has to be simplified to satisfy the third requirement of Section~\ref{sec:couplingSolidAndFluid}, concerning computation time performance.
As for fluid simulation in Section~\ref{sec:oceanWaves}, \framework assumes that the interaction from a solid to the fluid can be simulated as a 2D height field.
Furthermore, it considers a single frequency for the waves generated by a solid $M$.
These specific waves are called interactive waves.
These approximations may seem too strong, but as mentioned above, they are necessary to satisfy the requirements defined in Section~\ref{sec:couplingSolidAndFluid}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Grid translation}
\label{subsec:gridTranslation}
The waves generated by a given body $M$ are defined in its local coordinates system.
Hence, when $M$ moves, the waves have to move accordingly as done by Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
A first solution consists of applying the translation to the grid at times $t$ and $t-dt$ and then of using the integration scheme of Equation (\ref{eq:fdm_scheme}).
Nevertheless, these time-consuming shifts can be avoided by using the following domain translation:
\begin{equation}
	\begin{cases}
		k = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t)}{dx}\right\rfloor    \\
		l = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t)}{dx}\right\rfloor    \\
		o = & i - \left\lfloor\frac{\mathbf{p}_x(t+dt)-\mathbf{p}_x(t-dt)}{dx}\right\rfloor \\
		p = & j - \left\lfloor\frac{\mathbf{p}_z(t+dt)-\mathbf{p}_z(t-dt)}{dx}\right\rfloor
	\end{cases}
\end{equation}
where $\mathbf p(t)=\left[p_x(t)\ p_z(t)\right]^T$ is the position of the mesh $M$ at time $t$, $k$ and $l$ (respectively $o$ and $p$) are the shifted indices defined by the translation of $M$ at times $t+dt$ from times $t$ (respectively from times $t-dt$).
Then, the integration scheme can be rewritten as follows:
\begin{equation}
	\label{eq:fdm_scheme_corrected}
	h_{i,j}^{n+1} = d^n\left(a\left(h_{k+1,l}^n+h_{k-1,l}^n+h_{k,l+1}^n+h_{k,l-1}^n- 4 h_{k,l}^n\right) + 2 h_{k,l}^n - h_{o,p}^{n-1}\right)
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Evaluate the height of vertices in the mask}
\label{subsubsec:heightMask}
Knowing all the vertices in the mask, their height must be given as input of the FDM.
\Framework uses a simple model that tries to reproduce parts of the idea of Cords and Staadt (2009) \cite{cordsRealTimeOpenWater2009}.
By construction of the grid $Z$, $M$ has is bow oriented on the positive side of the $z$ axis, and its starboard on the positive side of the $x$ axis.
Hence, $h$ is made such that from the vertex at position $\mathbf x$, the front of the mask is above the free surface and the back below, in a way that is proportional to the speed of $M$.
Moreover, the boat forms a $V$ shape, with its center a bit below the extremity of the side.
